### 2014-05-15 brains-protocol.tex
Work on 3.3 and 4
* Further separate figure 3 into 3a and 3b
* Rework the presentation of binning of quotation families
* Rework 4 to be compatible with the above

We face an inference problem where, given all quotations and their occurrence timestamps, we should estimate which was the originating quotation for each instance of each quotation.
We therefore model the underlying quotation selection process by making a few additional assumptions which let us define quote-to-quote substitutions from the available data.

The main question in this problem concerns the manner in which we decide whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the dataset, \hbox{i.e.} from a source external to the data collection perimeter.
%consider later occurrences of a quotation that has already occurred earlier: the author of the later occurrences could have sourced an earlier occurrence and accurately copied it, they could have transformed an earlier occurrence of \emph{another} quote into this one, or they could have relied on information or other quotes from outside the dataset, i.e. from an external source we do not measure.

Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q_a$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q_b$) appears in other blogs on the 21st, 22nd and 23rd of January.
If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the first author of $q_b$ on the 21st based himself on $q_a$ as is shown in the top drawing of Figure~\ref{fig:substitution-temporal-binning}.
But what about the second and third occurrences of $q_b$, on the 22nd and 23rd?
Should we consider them to be substitutions based on $q_a$ %(i.e. re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) 
or accurate reproductions of the previous occurrences of $q_b$? \cam{là je distinguerais 3a [top] de 3b [middle \& bottom]} (options shown in Figure~\ref{fig:substitution-temporal-binning})?

\begin{figure}[h]
    \centering
    \def\svgwidth{\linewidth}
    \small
    \input{images/substitution-temporal-binning.pdf_tex}
    \caption{XXXXX}
    \label{fig:substitution-temporal-binning}
\end{figure}

To settle this question we bin the quote occurrences into fixed \emph{time bags} spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution
We further assume that when a quotation $q$ appears in time bag $n$, if it is not original (i.e. if not stemming from a source external to the dataset, e.g. initiating a new set of quotations), then it is based solely on the most frequent quotation $q_{max}$ in the preceding time bag ($n - 1$).
If $q$ differs from $q_{max}$ by only one word it is counted as a substitution from $q_{max}$ to~$q$.
In any other case, i.e. if $q$ and $q_{max}$ are the same or if $q$ and $q_{max}$ are different in other ways, the occurrence of $q$ is not considered to be an instance of substitution and is discarded.

This model has two main effects.
First, the binning influences how many times quote occurrences can be counted as substitutions: in the second drawing on Figure~\ref{fig:substitution-temporal-binning}, occurrences of $q_b$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
Second, the ``majority'' rule defines how quotes are sourced in substitutions: in the third drawing on Figure~\ref{fig:substitution-temporal-binning}, $q_b$ holds the majority and is the considered the basis for the last occurrence of $q_a$, in spite of $q_a$ having appeared earlier at the very beginning (indeed in the situation shown in Figure~\ref{fig:substitution-temporal-binning}, this seems the most likely scenario).

The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative models.
We identified and implemented eleven other such models, and they all yielded essentially the same results.
These models differ on four fundamental principles or constraints:

\begin{itemize}
    \item \textbf{Time} can be sliced into bins to build time bags (as in the model described above) or kept fine-grained by using sliding time bags,
    \item The \textbf{source time interval} from which a source quote is selected can span a fixed number of days, or can go back to the beginning of the quotation family,
    \item \textbf{Source quotes} in the source interval for substitutions can be further restricted, e.g. to only the most frequent quote in the interval as in the model described above,
    \item \textbf{Destination quotes} for substitutions can be further restricted, e.g. to only quotes that did not appear in the source time interval (ensuring it is the first occurrence of the destination quote).
\end{itemize}

These various flavours of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.

@@

To do so we use a variation of the model presented in the previous section: quote occurrences are binned into 2-day time bags, substitutions can stem from any quote in the time bag preceding the current one, but destination quotes must be absent from that preceding time bag. The last two points constitute a variation from the ``majority'' rule which let us detect substitution chains later on.

\rk{Adapt this last paragraph once the model is really chosen (see below)}

We then compute the distribution of word features for each time bag $\mathcal{B}_n$ of each quotation family, and sum those distributions over all quotation families.
This yields a distribution of feature values for each $n$, which is the simplest possible view of the state of an average quotation family in its $n$-th time bag, i.e. after $n \times \Delta t$ days.\footnote{For consistency, if we do this for $n$ going up to $N$, we only include quotation families that span long enough to have at least $N$ time bags.}
Such a computation makes no assumptions on the way quotations undergo substitutions over time.

@@

\bigskip
To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: using the above-mentioned substitution detection model, we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
The first time bag is therefore untouched, the second contains the quotations produced by substitutions from the first, the third by substitutions from the second, and so on.
As a result, the size of observed words drastically decreases across time, yet it unambiguously focuses on successive mutations.
\rk{add pedagogical figure}
We thus observe only the repeatedly-affected portion of the quotation families, and obtain a simple view of the longitudinal effect of cognitive bias, \hbox{i.e.} how quote features are evolving in the long term and, perhaps, converging towards specific attractors.

@@

>>>>

We face an inference problem where, given all quotations and their occurrence timestamps, we should estimate which was the originating quotation for each instance of each quotation.
We therefore model the underlying quotation selection process by making a few additional assumptions which let us define quote-to-quote substitutions from the available data.
The main issue at hand is deciding whether a later occurrence is a strict copy of an earlier occurrence, or a substitution of an even earlier occurrence, or perhaps even a substitution or copy from quotes appearing outside the dataset, \hbox{i.e.} from a source external to the data collection perimeter.

Let us give an example: say the quotation ``These accusations are false and \textbf{absurd}'' ($q_a$) appears in a blog on January 19, and the slightly different quotation ``These accusations are false and \textbf{incoherent}'' ($q_b$) appears in other blogs on the 21st, 22nd and 23rd of January.
If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the first author of $q_b$ on the 21st based himself on $q_a$ as is shown in Figure~\ref{fig:substitution-temporal-binning-a}.
But what about the second and third occurrences of $q_b$, on the 22nd and 23rd?
Should we consider them to be substitutions based on $q_a$ %(i.e. re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) 
or accurate reproductions of the previous occurrences of $q_b$? (Options shown in Figure~\ref{fig:substitution-temporal-binning-a}.)

\begin{figure}[h]
    \centering
    \subfloat[Possible paths from occurrence to occurrence]{
	    \def\svgwidth{\linewidth}
	    \small
	    \input{images/substitution-temporal-binning-a.pdf_tex}
	    \label{fig:substitution-temporal-binning-a}
	}
	\hfill \\
    \subfloat[Binned quotation family]{
	    \def\svgwidth{\linewidth}
	    \small
	    \input{images/substitution-temporal-binning-b.pdf_tex}
	    \label{fig:substitution-temporal-binning-b}
	}
	\caption{Temporal binning of quotation families}
    \label{fig:substitution-temporal-binning}
\end{figure}

To settle this question we bin the quote occurrences into fixed \emph{time bags} spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution.
Then when a quotation $q$ appears in time bag $n$, it is counted as a substitution from each quote $q^*$ in the preceding time bag ($n - 1$) from which it differs by only one word.
If no quote in the preceding time bag can qualify as a source in a substitution (i.e. $q$ differs from all the quotes in the preceding time bag by more than one word), the occurrence of $q$ is not considered to be an instance of substitution.
Such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q_b$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.

The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative substitution inferences.\footnote{In particular, time can be sliced into bins to build fixed time bags as is done here, or kept fine-grained by using sliding time bags.}
These various flavours of an ideal substitution detection model essentially change whether occurrences are considered as substitutions from another quote, repetitions of the original quote, or introduction of information external to the dataset.
We identified and implemented eleven other such models, and they all yielded essentially the same results.

@@

To do so we compute the distribution of word features for each time bag $\mathcal{B}_n$ of each quotation family, and sum those distributions over all quotation families.
This yields a distribution of feature values for each $n$, which is the simplest possible view of the state of an average quotation family in its $n$-th time bag, i.e. after $n \times \Delta t$ days.\footnote{For consistency, if we do this for $n$ going up to $N$, we only include quotation families that span long enough to have at least $N$ time bags.}
Such a computation, based solely on the binning of quotation families, makes no assumptions on the way quotations undergo substitutions over time.

@@

\bigskip
To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: using the substitution detection model described in the previous section, we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
The first time bag is therefore untouched, the second contains the quotations produced by substitutions from the first, the third by substitutions from the second, and so on.
As a result, the number of observed words drastically decreases across time, yet it unambiguously focuses on successive mutations.
\rk{add pedagogical figure}
We thus observe only the repeatedly-affected portion of the quotation families, and obtain a simple view of the longitudinal effect of cognitive bias, \hbox{i.e.} how quote features are evolving in the long term and, perhaps, converging towards specific attractors.

@@

\TB{This model has two main effects.
First, the binning influences how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-temporal-binning-b}, occurrences of $q_b$ on the 21st and 22nd are counted as substitutions, whereas the occurrence on the 23rd is not.
Second, the ``majority'' rule defines how quotes are sourced in substitutions: in the third drawing on Figure~\ref{fig:substitution-temporal-binning}, $q_b$ holds the majority and is the considered the basis for the last occurrence of $q_a$, in spite of $q_a$ having appeared earlier at the very beginning (indeed in the situation shown in Figure~\ref{fig:substitution-temporal-binning}, this seems to be the most likely scenario).}

\TB{\textbf{Source quotes} in the source interval for substitutions can be further restricted, e.g. to only the most frequent quote in the interval as in the model described above,}
